---
title: "Crop land cover by habitat use"
author: "Sydney Schmitter"
date: "6/30/2022"
output: 
  html_document:
    fig_width: 8
    fig_height: 14
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries
library(cdlTools)
library(raster)
library(tidyverse)
library(janitor)
library(sf)
library(stars)
library(ggspatial)
library(tmap)
library(units)
library(tabularaster)
library(osmdata)
library(here)
library(exactextractr)
library(CropScapeR)
library(foreign)
library(ggthemes)
library(knitr)

# Set ggplot2 themes
theme_set(theme_clean())
theme_update(
  plot.background = element_rect(color = NA),
  plot.title.position = "plot",
  plot.caption.position = "plot"
)

# Load CDL key
data("linkdata")

#setwd("~/Desktop/Github Repo/salmon_ag")
```
```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r load data, include=FALSE}
# Load habitat use recovery domain shape file
(sf_recoverydomain_use <- st_read(here("data/huc6_esu-withadds-ver5-alb27.shp")) %>% clean_names()) %>% names



# sf_recoverydomain <- st_read(here("data/recovery_subdomains/subdomains-ver7.shp"))
# Key variables are, with * as species code
# r_*: ESU name
# sp_*: ESU spawn/rear
# re_*: ESU rear/migrate
# mi_*: ESU migrate only

# Species codes are (first code is for r_*, second is for sp_*, re_*, mi_*)
# ch/l_chin: chinook
# ch_sp/l_chsp: spring chinook
# ch_sp_su/l_chsps: spring/summer chinook
# ch_fa/l_chfa: fall chinook
# ch_wi/l_chwi: winter chinook
# coho/l_coho: coho
# sthd/l_sthd: steelhead
# chum/l_chup: chum
# chum_su/l_chum: summer chum
# ch_su_fa: summer/fall chinook (no use variables)
# sock/l_sock: sockeye
# pink_*/l_pink: pink (oy/ey: odd year/even year; use data shared)


CDLwest = raster('cdl_west.tif') #this .tif is what we made down below, once you run the code below, it is stored for downstream use
#raster_cdl_merge <- raster("./data/cdl_west.tif")
  #projectRaster(raster_cdl_merge, crs=sf_recoverydomain_use) #This takes a long time
  #writeRaster(raster_cdl_merge, "./data-nogit/cdl_west_prof.tif", format = "GTiff", overwrite = TRUE) 
#raster_cdl_merge <- raster(here("./data-nogit/cdl_west_prof.tif"))
# raster_cdl_merge <- raster(here("/data/cdl_west.tif"))


```

## R Markdown

This document summarizes crop-specific land cover over the habitat range of listed salmonid populations on the West Coast, by habitat use types. Included are land cover summaries for 30 Distinct Population Segments (DPS). Details on each DPS life history are also reprinted from Crozier et al. (2019), for reference, along with their map of the eight distinct recover domains (ecoregions with distinct climatic and ecological characteristics).

```{r extract land cover, eval=FALSE, include=FALSE}
#This takes a while and is best left overnight to run!
#This should be a loop or lapply, but some data table manipulation that I'm very slow at is needed. Speed might be significantly improved with lapply

# fastDoCall <- function(what, args, quote = FALSE, envir = parent.frame()){
#   if (quote)
#     args <- lapply(args, enquote)
# 
#   if (is.null(names(args))){
#     argn <- args
#     args <- list()
#   }else{
#     # Add all the named arguments
#     argn <- lapply(names(args)[names(args) != ""], as.name)
#     names(argn) <- names(args)[names(args) != ""]
#     # Add the unnamed arguments
#     argn <- c(argn, args[names(args) == ""])
#     args <- args[names(args) != ""]
#   }
# 
#   if (class(what) == "character"){
#     if(is.character(what)){
#       fn <- strsplit(what, "[:]{2,3}")[[1]]
#       what <- if(length(fn)==1) {
#         get(fn[[1]], envir=envir, mode="function")
#       } else {
#         get(fn[[2]], envir=asNamespace(fn[[1]]), mode="function")
#       }
#     }
#     call <- as.call(c(list(what), argn))
#   }else if (class(what) == "function"){
#     f_name <- deparse(substitute(what))
#     call <- as.call(c(list(as.name(f_name)), argn))
#     args[[f_name]] <- what
#   }else if (class(what) == "name"){
#     call <- as.call(c(list(what, argn)))
#   }
# 
#   eval(call,
#        envir = args,
#        enclos = envir)
# }
# 
# 
# #We include an install check for the compiled data to avoid running again if not needed
# if(!file.exists("land_use_habitat_use_2021.RData")) {
# 
# speciesvector <- c("r_ch", "r_ch_sp", "r_ch_sp_su", "r_ch_fa", "r_ch_wi", "r_coho", "r_sthd", "r_chum", "r_chum_su", "r_sock", "r_pink_oy", "r_pink_ey")
# speciesselect <- c("sp_l_chin", "mi_l_chin", "sp_l_chsp", "mi_l_chsp", "sp_l_chsps", "mi_l_chsp", "sp_l_chfa",
#                    "mi_l_chfa", "sp_l_chwi", "re_l_chwi", "sp_l_coho", "mi_l_coho", "sp_l_sthd", "mi_l_sthd", "sp_l_chup", "mi_l_chup", "sp_l_chum", "mi_l_chum", "sp_l_sock","mi_l_sock","sp_l_pink","mi_l_pink","sp_l_pink", "mi_l_pink")
# 
# lapply(speciesvector, function(i))
#   sf_recoverydomain_use %>%
#   filter(!is.na(i)) %>%
#   select(c(objectid:states), i, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(i)) %>%
#              select(speciesselect[i]:speciesselect[i+1]) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration"))
#   )
# 
# #r_ch
# #pull species name
# species <- sf_recoverydomain_use %>%
#   tabyl(r_ch) %>% pull(1)
# 
# tibble("hab_use" = sf_recoverydomain_use %>%
#          st_drop_geometry() %>%
#          filter(r_ch == species[1]) %>%
#          select(sp_l_chin:mi_l_chin) %>%
#          as.matrix() %*% 1:3) %>% tabyl(hab_use)
# 
# #create habitat use column
# sf_recoverydomain_use_r_ch <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_ch)) %>%
#   select(c(objectid:states), r_ch, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_ch)) %>%
#              select(sp_l_chin:mi_l_chin) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration"))
#   )
# 
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_ch)
# #separates to just california coastal chinook habitat use
# ex <- exact_extract(CDLwest,b,include_cols = c("r_ch", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_ch <- ex  %>% rename(name = r_ch) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# write.csv(result_r_ch, "result_r_ch.csv")
# 
# #r_ch_sp
# sf_recoverydomain_use_r_ch_sp <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_ch_sp)) %>%
#   select(c(objectid:states), r_ch_sp, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_ch_sp)) %>%
#              select(sp_l_chsp:mi_l_chsp) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
#    #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_ch_sp)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_ch_sp", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_ch_sp <- ex  %>% rename(name = r_ch_sp) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# 
# #r_ch_sp_su
# sf_recoverydomain_use_r_ch_sp_su <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_ch_sp_su)) %>%
#   select(c(objectid:states), r_ch_sp_su, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_ch_sp_su)) %>%
#              select(sp_l_chsps:mi_l_chsps) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_ch_sp_su)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_ch_sp_su", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_ch_sp_su <- ex  %>% rename(name = r_ch_sp_su) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# #r_ch_fa
# sf_recoverydomain_use_r_ch_fa <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_ch_fa)) %>%
#   select(c(objectid:states), r_ch_fa, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_ch_fa)) %>%
#              select(sp_l_chfa:mi_l_chfa) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_ch_fa)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_ch_fa", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_ch_fa <- ex  %>% rename(name = r_ch_fa) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# #r_ch_wi
# sf_recoverydomain_use_r_ch_wi <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_ch_wi)) %>%
#   select(c(objectid:states), r_ch_wi, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_ch_wi)) %>%
#              select(sp_l_chwi:re_l_chwi) %>%
#              as.matrix() %*% 1:2)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2), labels = c("spawn+rearing", "rearing+migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_ch_wi)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_ch_wi", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_ch_wi <- ex  %>% rename(name = r_ch_wi) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# #r_coho
# sf_recoverydomain_use_r_coho <-
# sf_recoverydomain_use %>%
#   filter(!is.na(r_coho)) %>%
#   select(c(objectid:states), r_coho, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_coho)) %>%
#              select(sp_l_coho:mi_l_coho) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_coho)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_coho", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_coho <- ex  %>% rename(name = r_coho) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# #r_sthd
# sf_recoverydomain_use_r_sthd <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_sthd)) %>%
#   select(c(objectid:states), r_sthd, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_sthd)) %>%
#              select(sp_l_sthd:mi_l_sthd) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_sthd)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_sthd", "hab_use"))
# 
# ex1 <- ex[1:2606]
# ex1 <- fastDoCall(rbind,ex1)
# 
# result_r_sthd1 <- ex1 %>% rename(name = r_sthd) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# write.csv(result_r_sthd1, file = "result_r_sthd1")
# 
# ex2 <- ex[2606:3606]
# ex2 <- fastDoCall(rbind,ex2)
# 
# result_r_sthd2 <- ex2 %>% rename(name = r_sthd) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# write.csv(result_r_sthd2, file = "result_r_sthd2")
# 
# 
# ex3 <- ex[3606:4606]
# ex3 <- fastDoCall(rbind, ex3)
# 
# result_r_sthd3 <- ex3 %>% rename(name = r_sthd) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# write.csv(result_r_sthd3, file = "result_r_sthd3")
# 
# 
# ex4 <- ex[4606:5211]
# ex4 <- fastDoCall(rbind, ex4)
# 
# result_r_sthd4 <- ex4 %>% rename(name = r_sthd) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# write.csv(result_r_sthd4, file = "result_r_sthd4")
# 
# 
# 
# #r_chum
# sf_recoverydomain_use_r_chum <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_chum)) %>%
#   select(c(objectid:states), r_chum, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_chum)) %>%
#              select(sp_l_chup:mi_l_chup) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_chum)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_chum", "hab_use"))
# split(x = ex, f =)
# ex <- fastDoCall(rbind,ex)
# result_r_chum <- ex %>% rename(name = r_chum) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# #r_chum_su
# sf_recoverydomain_use_r_chum_su <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_chum_su)) %>%
#   select(c(objectid:states), r_chum_su, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_chum_su)) %>%
#              select(sp_l_chum:mi_l_chum) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_chum_su)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_chum_su", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_chum_su <- ex %>% rename(name = r_chum_su) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# 
# #r_sock
# sf_recoverydomain_use_r_sock <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_sock)) %>%
#   select(c(objectid:states), r_sock, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_sock)) %>%
#              select(sp_l_sock:mi_l_sock) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_sock)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_sock", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_sock <- ex %>% rename(name = r_sock) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# #r_pink_oy
# sf_recoverydomain_use_r_pink_oy <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_pink_oy)) %>%
#   select(c(objectid:states), r_pink_oy, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_pink_oy)) %>%
#              select(sp_l_pink:mi_l_pink) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_pink_oy)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_pink_oy", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_pink_oy <- ex %>% rename(name = r_pink_oy) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# #r_pink_ey
# sf_recoverydomain_use_r_pink_ey <-
#   sf_recoverydomain_use %>%
#   filter(!is.na(r_pink_ey)) %>%
#   select(c(objectid:states), r_pink_ey, c(regime01b:r_ch03_c), geometry) %>%
#   bind_cols(
#     tibble("hab_use" = sf_recoverydomain_use %>%
#              st_drop_geometry() %>%
#              filter(!is.na(r_pink_ey)) %>%
#              select(sp_l_pink:mi_l_pink) %>%
#              as.matrix() %*% 1:3)
#   ) %>%
#   mutate(
#     hab_use = case_when(hab_use == 0 ~ NA_real_, TRUE ~ hab_use),
#     hab_use = factor(hab_use, levels = c(1, 2, 3), labels = c("spawn+rearing", "rearing+migration", "migration")),
#   )
# #extract and create summary table
# b <- st_as_sf(sf_recoverydomain_use_r_pink_ey)
# ex <- exact_extract(CDLwest,b,include_cols = c("r_pink_ey", "hab_use"))
# ex <- fastDoCall(rbind,ex)
# result_r_pink_ey <- ex %>% rename(name = r_pink_ey) %>% group_by(name, hab_use) %>% summarise(
#   land.use = sort(unique(value)),
#   count = table(value),
#   freq = table(value) / length(value)
# )
# 
# result_r_ch_fa <- read.csv('result_r_ch_fa')
# result_r_ch <- read.csv('result_r_ch')
# result_r_ch_sp <- read.csv('result_r_ch_sp')
# result_r_ch_sp_su <- read.csv('result_r_ch_sp_su')
# result_r_ch_wi <- read.csv('result_r_ch_wi')
# result_r_chum <- read.csv('result_r_chum')
# result_r_chum_su <- read.csv('result_r_chum_su')
# result_r_coho <- read.csv('result_r_coho')
# result_r_pink_ey <- read.csv('result_r_pink_ey')
# result_r_pink_oy <- read.csv('result_r_pink_oy')
# result_r_sock <- read.csv('result_r_sock')
# result_r_sthd1 <- read.csv('result_r_sthd1')
# result_r_sthd2 <- read.csv('result_r_sthd2')
# result_r_sthd3 <- read.csv('result_r_sthd3')
# result_r_sthd4 <- read.csv('result_r_sthd4')
# result_r_sthd <- rbind(result_r_sthd1, result_r_sthd2, result_r_sthd3, result_r_sthd4)
# 
# 
# land_use_habitat_use <- rbind(result_r_ch, result_r_ch_fa, result_r_ch_sp, result_r_sthd, result_r_ch_sp_su, result_r_ch_wi, result_r_chum, result_r_chum_su, result_r_coho, result_r_pink_ey, result_r_pink_oy, result_r_sock)
# 
# linkdata <- linkdata %>%
#   mutate(cropgroup = if_else(str_detect(Crop, "Barley|Rye|Wheat|Oats|Triticale|Sorghum|Buckwheat|Millet|Dbl|Other_Small_Grains|Soybeans|Speltz|Other_Small_Grains"), "smgrains",
#                            if_else(str_detect(Crop, "Walnuts|Almonds|Pecans|Pistachios"), "nuts" ,
#                                    if_else(str_detect(Crop, "Apples|Apricots|berries|Cantaloupes|Cherries|Citrus|Watermelons|Melons|Nectarines|Oranges|Peaches|Pears|Plums|Pomegranates|Other_Tree|Olives|Prunes"), "fruits",
#                                            if_else(str_detect(Crop, "Tomatoes|Asparagus|Broccoli|Cabbage|Carrots|Cauliflower|Cucumbers|Garlic|Gourds|Greens|Lettuce|Onions|Peas|Peppers|Popcorn|Pumpkins|Radish|Squash|Sweet_Corn|Sweet_Potatoes|Turnips|Misc_Vegs|Misc_Vegs_&_Fruits|Celery|Eggplants"), "vegetables",
#                                                    if_else(str_detect(Crop,  "Chick_Peas|Christmas_Trees|Dry_Beans|Flaxseed|Herbs|Hops|Lentils|Mint|Mustard|Other_Crops|Rape|Safflower|Sod/Grass_Seed|Sugarbeets|Sunflower|Vetch|Pop_or_Orn_Corn|Peanuts|Tobacco|Canola|Camelina|Sugarcane|Clover/Wildflowers"), "other_crops",
#                                                            if_else(str_detect(Crop,  "Alfalfa|Other Hay|Switchgrass"), "hay",
#                                                                    if_else(str_detect(Crop,  "Rice"), "rice",
#                                                                            if_else(str_detect(Crop,  "Cotton"), "cotton",
#                                                                                    if_else(str_detect(Crop,  "Fallow"), "fallow",
#                                                                                            if_else(str_detect(Crop,  "Grapes"), "grapes",
#                                                                                                    if_else(str_detect(Crop,  "Pasture"), "pasture",
#                                                                                                            "other"))))))))))))
# linkdata$cropgroup <- if_else(linkdata$Crop == "Potatoes", "potatoes", linkdata$cropgroup)
# linkdata$cropgroup <- if_else(linkdata$Crop == "Corn", "corn", linkdata$cropgroup)
# linkdata$cropgroup <- if_else(str_detect(linkdata$Crop,"Open Water|Water"), "Water", linkdata$cropgroup)
# 
# 
# land_use_habitat_use <- left_join(land_use_habitat_use, linkdata, by = c('land.use' = 'MasterCat'))
# land_use_habitat_use <- land_use_habitat_use  %>%
# mutate(
#     spawning = case_when(hab_use == "spawn+rearing" ~ 1),
#     rearing = case_when(hab_use == "spawn+rearing" | hab_use == "rearing+migration" ~ 1),
#     migration = case_when(hab_use == "rearing+migration" | hab_use == "migration" ~ 1))
# 
# save(land_use_habitat_use, file = here("land_use_habitat_use_2021.RData")) #save data for use without running extract code, which is slow
# write.dta(land_use_habitat_use, here("./output/land_use_habitat_use_2021.dta")) #save data as Stata file for Lisa
# 
# }

```



```{r figure, echo=FALSE}
# setwd(wd)
load(here("land_use_habitat_use_2021.RData")) #load data if not already loaded

land_use_habitat_use_long <- land_use_habitat_use %>% pivot_longer(cols = c("spawning", "rearing", "migration"), names_to = "use", values_to = "code", values_drop_na = TRUE)

# Build land cover bar plot
p_cover <-
  land_use_habitat_use_long %>%
  # Isolate just the one species
#  filter(name == species[2]) %>%
  filter(cropgroup != "Water") %>%
  filter(cropgroup != "other") %>%
  # Exlude pasture if desired
#  filter(cropgroup != "pasture") %>%  
  filter(!is.na(use)) %>%
  ggplot() +
  # Plot the land cover breakdown
  geom_col(aes(x = use, y = freq, fill = cropgroup), position = "stack") +
   coord_flip() +
   scale_fill_manual(
     "Land cover (CDL2018)",
      limits = c( "corn", "cotton", "fallow", "fruits", "grapes","hay","other_crops", "nuts",  "pasture", "potatoes", "rice", "smgrains",  "vegetables"),
      labels = c("corn", "cotton", "fallow", "fruits", "grapes","hay","other crops", "nuts", "pasture", "potatoes", "rice", "small grains",  "vegetables"),
      values = c( "darkgoldenrod", "goldenrod", "grey", "darkred", "purple", "lightgreen", "gold", "tan", "darkgreen", "black", "violet","pink",  "springgreen3")) +
  # Set up the y-axis
  scale_y_continuous(breaks = c(0, 0.1,  0.2,  0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1), labels = scales::label_percent(), limits = c(0, 1)) +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
  ) +
  xlab('Habitat use') + ylab('Percent of habitat use covered by crop') +
  facet_wrap(~name, ncol = 2) +
  ggtitle("Agricultural land cover by ESU and habitat use")
plot(p_cover)

```

The same figure but excluding pasture: 

```{r figure2, echo=FALSE}
# setwd(wd)
# load("./output/land_use_habitat_use.RData") #load data if not already loaded


# Build land cover bar plot
p_cover_nopasture <-
  land_use_habitat_use_long %>%
  # Isolate just the one species
#  filter(name == species[2]) %>%
  filter(cropgroup != "Water") %>%
  filter(cropgroup != "other") %>%
  filter(cropgroup != "pasture") %>%  
  filter(!is.na(use)) %>%
  ggplot() +
  # Plot the land cover breakdown
  geom_col(aes(x = use, y = freq, fill = cropgroup), position = "stack") +
   coord_flip() +
   scale_fill_manual(
     "Land cover (CDL2018)",
      limits = c( "corn", "cotton", "fallow", "fruits", "grapes","hay","other_crops", "nuts",  "potatoes", "rice", "smgrains",  "vegetables"),
      labels = c("corn", "cotton", "fallow", "fruits", "grapes","hay","other crops", "nuts",  "potatoes", "rice", "small grains",  "vegetables"),
      values = c( "darkgoldenrod", "goldenrod", "grey", "darkred", "purple", "lightgreen", "gold", "tan",  "black", "violet","pink",  "springgreen3")) +
  # Set up the y-axis
  scale_y_continuous(breaks = c(0, 0.1,  0.2,  0.3, 0.4, 0.5, 0.6), labels = scales::label_percent(), limits = c(0, .6)) +
  theme(
    legend.position = "bottom",
    legend.box = "horizontal",
  ) +
  xlab('Habitat use') + ylab('Percent of habitat use covered by crop') +
  facet_wrap(~name, ncol = 2) +
  ggtitle("Agricultural land cover by ESU and habitat use, pasture excluded")
plot(p_cover_nopasture)

```

Notice that in the Upper Willamette River Chinook Salmon figure, a large amount of crop coverage is "other crops", which is Sod/Grass seed.


```{r detail, echo=FALSE, results = 'asis'}
# UW_chinook <- 
#   land_use_habitat_use %>%
#   ungroup() %>%
#   filter(name == "Upper Willamette River Chinook Salmon") %>%
#   filter(cropgroup == "other_crops") %>%
#   filter(!is.na(hab_use)) %>%
#   select(-name, -cropgroup, -land.use, -count) %>%
#   rowwise() %>%
#   arrange(hab_use, -as.numeric(freq)) %>%
#   mutate(freq = scales::percent(as.numeric(freq), 0.11)) %>%
#   mutate(freq = case_when(freq == "0.0%" ~ "< 0.1%", TRUE ~ freq)) %>%
#   mutate(hab_use = case_when(hab_use == "spawn+rearing" ~ "Spawning & rearing", TRUE ~ "Rearing & migration")) %>%
#   select("Habitat use" = hab_use, Crop, "% land cover, CDL2018" = freq)
#           
# kable(UW_chinook, caption = 'List of "other crops" in Upper Willamette River Chinook ESU range') %>%
#   kableExtra::kable_styling() %>%
#   kableExtra::collapse_rows(1, valign = "top")
```
