---
title: "recreation_salmon_agmaps_loop"
author: "Sydney Schmitter"
date: "6/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::knit_meta(class=NULL, clean = TRUE)

# Clear existing environment
rm(list = ls())

# Load libraries
library(cdlTools)
library(raster)
library(tidyverse)
library(ggthemes)
library(janitor)
library(sf)
library(stars)
library(ggspatial)
library(tmap)
library(units)
library(tabularaster)
library(cowplot)
library(knitr)
library(kableExtra)
library(here)



# Set ggplot2 themes
theme_set(theme_clean())
theme_update(
  plot.background = element_rect(color = NA),
  plot.title.position = "plot",
  plot.caption.position = "plot"
)

setwd("~/Desktop/Github Repo/salmon_ag")
```

## Agricultural land use in salmon habitat

This document summarizes land cover over the habitat range of listed salmonid populations on the West Coast, with a focus on agriculture. Included are land cover summaries for 30 Distinct Population Segments (DPS).

```{r data-prep, echo=FALSE, message=FALSE, include=FALSE}
# Get CDL data, process outlined on recreationsalmonag.Rmd
CDLtif = raster('cdl_west.tif')
CDLwest <- CDLtif

#clean up recovery domain names, simplify naming of targeted species to summarize
sf_recoverydomain <- st_read(here("data/recovery_subdomains/subdomains-ver7.shp")) %>% clean_names()
sf_recoverydomain <-
  sf_recoverydomain %>%
  distinct(.keep_all = TRUE) %>%
  dplyr::filter(`class` == "Accessible") %>%
  select(-class, -sourcethm) %>%
  mutate(
    species2 = case_when(
      species == "Steelhead" ~ "Steelhead",
      str_detect(species, "Chinook") ~ "Chinook",
      str_detect(species, "Chum") ~ "Chum",
      str_detect(species, "Pink") ~ "Pink",
      TRUE ~ species
    ),
    species3 = case_when(
      species == "Steelhead" ~ "Trout",
      TRUE ~ "Salmon"
    ),
    domain = case_when(
      subdomain == "Washington Coast" ~ "Puget Sound",
      subdomain == "Snake River" ~ "Interior Columbia",
      subdomain == "Middle Columbia River" ~ "Interior Columbia",
      subdomain == "Upper Columbia River" ~ "Interior Columbia",
      subdomain == "Lower Columbia River" ~ "Willamette/Lower Columbia",
      subdomain == "Upper Willamette River" ~ "Willamette/Lower Columbia",
      subdomain == "California Central Valley" ~ "Central Valley",
      subdomain == "North-Central California Coast" ~ "North-Central California Coast",
      TRUE ~ subdomain
    ),
    area = st_area(geometry),
    area = set_units(area, km^2)
  ) %>%
  st_transform(CDLwest@crs)

#Need cleaned csv
df_crdcount <- read_csv(here("data/data_subdomains_crd.csv"))

# Load states borders and transform to match crs
#create data frame for CA, ID, OR, WA
states_core <- 
  st_as_sf(maps::map("state", regions = c("california", "oregon", "washington", "idaho"), plot = FALSE, fill = TRUE)) %>%
  st_transform(st_crs(sf_recoverydomain))
#create data frame for western half of the united states
states_expand <-
  st_as_sf(
    maps::map(
      "state", 
      regions = 
        c(
          "california", 
          "oregon", 
          "washington", 
          "idaho", 
          "montana", 
          "wyoming",
          "arizona",
          "nevada",
          "utah",
          "colorado",
          "new mexico"
        ), plot = FALSE, fill = TRUE)
  ) %>% st_transform(st_crs(sf_recoverydomain))
#create data frame excluding USA? but keeping mexico + canada
land <- st_as_sf(maps::map("world", regions = c("Canada", "Mexico"), plot = FALSE, fill = TRUE)) %>% st_transform(st_crs(sf_recoverydomain))


```


```{r lapply, echo=FALSE}
# List of distinct population segments
dps <- df_crdcount %>% distinct(esu_dps) %>% arrange(esu_dps) %>% pull(esu_dps)
#creat slice of dps
dps <- dps[1:37] #should be 1:37 but my computer can't hold them all in memory

invisible(
  for(i in dps) {
           cat("  \n###", i)
           cat("  \n**Status:**", df_crdcount %>% filter(esu_dps == i) %>% pull(status))
           cat("  \n**Domain:**", df_crdcount %>% filter(esu_dps == i) %>% pull(domain))
           cat("  \n")
           
           #Plot the agricultural land uses inside each area
           border <- sf_recoverydomain %>% 
             arrange(area) %>% 
             filter(esu_dps == i)
           
           clip1 <- raster::crop(CDLwest, extent(border)) # Clip cdl to rectangle extents of the polygon
           clip2 <- mask(clip1, border) # Mask cdl to only what's within polygon
           
           
           df_clip_pts <- rasterToPoints(clip2, spatial = TRUE)
           df_clip2 <- as.data.frame(df_clip_pts, xy = TRUE) %>% drop_na(cdl_west)
           rm(df_clip_pts)
           
          types<- df_clip2  %>%
              group_by(cdl_west)  %>%
              count
        
          types<-types %>%
            mutate(cropgroup=if_else(str_detect(cdl_west, "21|27|22|23|24|28|205|4|39|29|254|241|240|239|238|237|236|235|234|233|232|231|230|228|226|225"), "Small Grains",
                           if_else(str_detect(cdl_west, "76|75|74|204"), "Nuts" ,
                                   if_else(str_detect(cdl_west, "68|223|55|221|242|250 209|66|72|48|213|218|212|67|77|220|217|71|211"), "Fruits",
                                           if_else(str_detect(cdl_west, "54|207|214|243|206|244|50|208|249|219|227|49|53|216|13|229|246|222|12|46|247"), "Vegetables",
                                                   if_else(str_detect(cdl_west,  "51|70|42|32|57|56|52|14|47|35|44|33|59|41|6|224"), "Other Crops",                
                                                           if_else(str_detect(cdl_west,  "36|37"), "Hay",
                                                                   if_else(str_detect(cdl_west,  "3"), "Rice",
                                                                           if_else(str_detect(cdl_west,  "2"), "Cotton",
                                                                                   if_else(str_detect(cdl_west,  "61"), "Fallow",
                                                                                           if_else(str_detect(cdl_west,  "69"), "Grapes",
                                                                                                   if_else(str_detect(cdl_west,  "62"), "Pasture",
                                                                                                           "Other"))))))))))))
types$cropgroup <- if_else(types$cdl_west == "43", "Potatoes", types$cropgroup)
types$cropgroup <- if_else(types$cdl_west == "1", "Corn", types$cropgroup)
types$cropgroup <- if_else(types$cdl_west == "111", "Open Water", types$cropgroup)
           
          # Join types to the land use data frame
           df_clip2_types = left_join(df_clip2, types)
           # Plot crop categories (types)
           p_ag_cover <-
             ggplot() +
             geom_raster(data = df_clip2_types, aes(x = x, y = y, fill = cropgroup)) +
             scale_fill_manual(
                        "Land cover (CDL2021)",
      limits = c( "corn", "cotton", "fallow", "fruits", "grapes","hay","other_crops", "nuts",  "pasture", "potatoes", "rice", "smgrains",  "vegetables", "Open Water", "other"),
      labels = c("corn", "cotton", "fallow", "fruits", "grapes","hay","other crops", "nuts", "pasture", "potatoes", "rice", "small grains",  "vegetables", "Open Water", "other"),
      values = c( "darkgoldenrod", "goldenrod", "grey", "darkred", "purple", "lightgreen", "gold", "tan", "darkgreen", "black", "violet","pink",  "springgreen3", "darkblue", "antiquewhite1")) +
  
             ggtitle(i) +
             theme_map()
           plot(p_ag_cover)
         }
  )
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
