---
title: "Agricultural Land Cover and Pacific Salmonid Habitat"
author: "B. Van Deynze"
date: "July 7, 2020"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
---
```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
 $(document).ready(function() {
    $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
    // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```

```{r setup, include=FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
# knitr::opts_chunk$set(fig.dim = c(8, 6))

# Clear existing environment
rm(list = ls())

# Load libraries
library(cdlTools)
library(raster)
library(tidyverse)
library(ggthemes)
library(janitor)
library(sf)
library(stars)
library(ggspatial)
library(tmap)
library(units)
library(tabularaster)
library(cowplot)
library(knitr)
library(kableExtra)
library(here)

# Set ggplot2 themes
theme_set(theme_clean())
theme_update(
  plot.background = element_rect(color = NA),
  plot.title.position = "plot",
  plot.caption.position = "plot"
)

# Set top-level working directory
# wd <- "C:/Users/braeden.vandeynze/Documents/Salmon Cost Project/Analysis/"  # For Braeden; comment out for other users
# wd <- "C:/Users/Braeden/Desktop/NOAA/Analysis"
# wd <- "D:/Braeden/OneDrive/Documents/My Work/NOAA/Agriculture/Analysis"
# knitr::opts_knit$set(root.dir = wd)
wd <- "C:/Users/Lisa.Pfeiffer/Documents/GitHub/salmon_ag/"
 setwd(wd)
# setwd("./data")
```


```{r data-prep, echo=FALSE, message=FALSE, include=FALSE}
# Get CDL data
# raster_cdl <-
#   getCDL(
#     c("CA", "WA", "OR", "ID"),
#     2017
#   )
# NAvalue(raster_cdl$CA2017) <- 0
# NAvalue(raster_cdl$OR2017) <- 0
# NAvalue(raster_cdl$WA2017) <- 0
# NAvalue(raster_cdl$ID2017) <- 0
# raster_cdl_merge <- raster::merge(raster_cdl$CA2017, raster_cdl$OR2017, raster_cdl$WA2017, raster_cdl$ID2017, overlap = TRUE)
# plot(raster_cdl_merge)
# writeRaster(raster_cdl_merge, "cdl_west.tif", format = "GTiff", overwrite = TRUE) # Saves data for downstream use
raster_cdl_merge <- raster(paste0(here("data"), "/cdl_west.tif"))

sf_recoverydomain <- st_read(paste0(here("data"), "/recovery_subdomains/subdomains-ver7.shp")) %>% clean_names()
sf_recoverydomain <-
  sf_recoverydomain %>%
  distinct(.keep_all = TRUE) %>%
  dplyr::filter(`class` == "Accessible") %>%
  select(-class, -sourcethm) %>%
  mutate(
    species2 = case_when(
      species == "Steelhead" ~ "Steelhead",
      str_detect(species, "Chinook") ~ "Chinook",
      str_detect(species, "Chum") ~ "Chum",
      str_detect(species, "Pink") ~ "Pink",
      TRUE ~ species
    ),
    species3 = case_when(
      species == "Steelhead" ~ "Trout",
      TRUE ~ "Salmon"
    ),
    domain = case_when(
      subdomain == "Washington Coast" ~ "Puget Sound",
      subdomain == "Snake River" ~ "Interior Columbia",
      subdomain == "Middle Columbia River" ~ "Interior Columbia",
      subdomain == "Upper Columbia River" ~ "Interior Columbia",
      subdomain == "Lower Columbia River" ~ "Willamette/Lower Columbia",
      subdomain == "Upper Willamette River" ~ "Willamette/Lower Columbia",
      subdomain == "California Central Valley" ~ "Central Valley",
      subdomain == "North-Central California Coast" ~ "North-Central California Coast",
      TRUE ~ subdomain
    ),
    area = st_area(geometry),
    area = set_units(area, km^2)
  ) %>%
  st_transform(raster_cdl_merge@crs)

df_crdcount <- read_csv(paste0(here("data"), "/data_subdomains_crd.csv"))

# Add collapsed categories
df_crdcount_summ <-
  df_crdcount %>%
  rowwise() %>%
  mutate(
    v_summ_count_total = sum(c_across(contains("_count_")), na.rm = TRUE),
    v_summ_count_developed = sum(c_across(contains("developed") &
                                            contains("_count_")), na.rm = TRUE),
    v_summ_pct_developed = sum(c_across(contains("developed") &
                                          contains("_pct_")), na.rm = TRUE),
    v_summ_count_forest = sum(c_across(contains("forest") &
                                         contains("_count_")), na.rm = TRUE),
    v_summ_pct_forest = sum(c_across(contains("forest") &
                                       contains("_pct_")), na.rm = TRUE),
    v_summ_count_smgrains = sum(c_across((
      contains("rye") |
        contains("barley") |
        contains("winter_wheat") |
        contains("oats") |
        contains("sprint_wheat") |
        contains("triticale") |
        contains("sorghum") |
        contains("buckwheat") |
        contains("millet") |
        contains("durum_wheat") |
        contains("dbl_crop")
    ) & contains("_count_")
    ), na.rm = TRUE),
    v_summ_pct_smgrains = sum(c_across((
      contains("rye") |
        contains("barley") |
        contains("winter_wheat") |
        contains("oats") |
        contains("sprint_wheat") |
        contains("triticale") |
        contains("sorghum") |
        contains("buckwheat") |
        contains("millet") |
        contains("durum_wheat") |
        contains("dbl_crop")
    )  & contains("_pct_")
    ), na.rm = TRUE),
    v_summ_count_nuts = sum(c_across((
      contains("walnuts") |
        contains("almonds") |
        contains("pecans") |
        contains("pistachios")
    ) & contains("_count_")
    ), na.rm = TRUE),
    v_summ_pct_nuts = sum(c_across((
      contains("walnuts") |
        contains("almonds") |
        contains("pecans") |
        contains("pistachios")
    ) & contains("_pct_")
    ), na.rm = TRUE),
    v_summ_count_fruit = sum(c_across((
      contains("apples") |
        contains("apricots") |
        contains("blueberries") |
        contains("caneberries") |
        contains("cantaloupes") |
        contains("cherries") |
        contains("citrus") |
        contains("honeydew_melons") |
        contains("nectarines") |
        contains("oranges") |
        contains("peaches") |
        contains("pears") |
        contains("plums") |
        contains("pomegranates") |
        contains("strawberries") |
        contains("other_tree_crops") |
        contains("olives") |
        contains("watermelons")
    ) & contains("_count_")
    ), na.rm = TRUE),
    v_summ_pct_fruit = sum(c_across((
      contains("apples") |
        contains("apricots") |
        contains("blueberries") |
        contains("caneberries") |
        contains("cantaloupes") |
        contains("cherries") |
        contains("citrus") |
        contains("honeydew_melons") |
        contains("nectarines") |
        contains("oranges") |
        contains("peaches") |
        contains("pears") |
        contains("plums") |
        contains("pomegranates") |
        contains("strawberries") |
        contains("other_tree_crops") |
        contains("olives") |
        contains("watermelons")
    ) & contains("_pct_")
    ), na.rm = TRUE),
    v_summ_count_veg = sum(c_across((
      contains("tomatoes") |
        contains("asparagus") |
        contains("broccoli") |
        contains("cabbage") |
        contains("carrots") |
        contains("cauliflower") |
        contains("cucumbers") |
        contains("garlic") |
        contains("gourds") |
        contains("greens") |
        contains("lettuce") |
        contains("onions") |
        contains("peas") |
        contains("peppers") |
        contains("pop_or_orn_corn") |
        contains("pumpkins") |
        contains("radishes") |
        contains("squash") |
        contains("sweet_corn") |
        contains("sweet_potatoes") |
        contains("turnips")
    ) & contains("_count_")
    ), na.rm = TRUE),
    v_summ_pct_veg = sum(c_across((
      contains("tomatoes") |
        contains("asparagus") |
        contains("broccoli") |
        contains("cabbage") |
        contains("carrots") |
        contains("cauliflower") |
        contains("cucumbers") |
        contains("garlic") |
        contains("gourds") |
        contains("greens") |
        contains("lettuce") |
        contains("onions") |
        contains("peas") |
        contains("peppers") |
        contains("pop_or_orn_corn") |
        contains("pumpkins") |
        contains("radishes") |
        contains("squash") |
        contains("sweet_corn") |
        contains("sweet_potatoes") |
        contains("turnips")
    ) & contains("_pct_")
    ), na.rm = TRUE),
    v_summ_count_othercrops = sum(c_across((
      contains("canola") |
        contains("chick_peas") |
        contains("christmas_trees") |
        contains("dry_beans") |
        contains("flaxseed") |
        contains("herbs") |
        contains("hops") |
        contains("lentils") |
        contains("mint") |
        contains("misc_vegs_fruits") |
        contains("mustard") |
        contains("other_crops") |
        contains("rape_seed") |
        contains("safflower") |
        contains("sod_grass_seed") |
        contains("sugarbeets") |
        contains("sunflower") |
        contains("vetch")
    ) & contains("_count_")
    ), na.rm = TRUE),
    v_summ_pct_othercrops = sum(c_across((
      contains("canola") |
        contains("chick_peas") |
        contains("christmas_trees") |
        contains("dry_beans") |
        contains("flaxseed") |
        contains("herbs") |
        contains("hops") |
        contains("lentils") |
        contains("mint") |
        contains("misc_vegs_fruits") |
        contains("mustard") |
        contains("other_crops") |
        contains("rape_seed") |
        contains("safflower") |
        contains("sod_grass_seed") |
        contains("sugarbeets") |
        contains("sunflower") |
        contains("vetch")
    ) & contains("_pct_")
    ), na.rm = TRUE),
    v_summ_count_hay = sum(c_across((
      contains("alfalfa") |
        contains("other_hay_non_alfalfa")
    ) & contains("_count_")
    ), na.rm = TRUE),
    v_summ_pct_hay = sum(c_across((
      contains("alfalfa") |
        contains("other_hay_non_alfalfa")
    ) & contains("_pct_")
    ), na.rm = TRUE),
    v_summ_count_pasture = sum(c_across(contains("pasture") &
                                          contains("_count_")), na.rm = TRUE),
    v_summ_pct_pasture = sum(c_across(contains("pasture") &
                                        contains("_pct_")), na.rm = TRUE),
    v_summ_count_ag = sum(c_across((
      !contains("wetands") &
        !contains("developed") &
        !contains("forest") &
        !contains("clover_wildflowers") &
        !contains("aquaculture") &
        !contains("ice_snow") &
        !contains("background") &
        !contains("barren") &
        !contains("shrubland") &
        !contains("water") &
        !contains("summ")
    ) & contains("_count_")
    ), na.rm = TRUE),
    v_summ_pct_ag = sum(c_across((
      !contains("wetlands") &
        !contains("developed") &
        !contains("forest") &
        !contains("clover_wildflowers") &
        !contains("aquaculture") &
        !contains("ice_snow") &
        !contains("background") &
        !contains("barren") &
        !contains("shrubland") &
        !contains("water") &
        !contains("summ")
    ) & contains("_pct_")
    ), na.rm = TRUE),
    
    v_summ_pct_corn = sum(c_across(
        matches("v_pct_corn")), na.rm = TRUE),
    v_summ_pct_cotton = sum(c_across(
      matches("v_pct_cotton")), na.rm = TRUE),
    v_summ_pct_grapes = sum(c_across(
      matches("v_pct_grapes")), na.rm = TRUE),
    v_summ_pct_rice = sum(c_across(
      matches("v_pct_rice")), na.rm = TRUE),
    v_summ_pct_fallow = sum(c_across(
      matches("v_pct_fallow_idle_cropland")), na.rm = TRUE),
    v_summ_pct_potatoes = sum(c_across(
      matches("v_pct_potatoes")), na.rm = TRUE),
   
    v_summ_count_ag_nopasture = v_summ_count_ag - v_summ_count_pasture,
    v_summ_pct_ag_nopasture = v_summ_pct_ag - v_summ_pct_pasture,
    v_summ_pct_shrubland = v_pct_shrubland,
    v_summ_count_other = v_summ_count_total - v_summ_count_ag + v_summ_count_developed + v_summ_count_forest,
    v_summ_pct_other = 1 - v_summ_pct_ag - v_summ_pct_developed - v_summ_pct_forest - v_summ_pct_shrubland
  )

# Transform for plotting
df_crdcount_long <-
  df_crdcount %>%
  select(-starts_with("v_pct")) %>%
  pivot_longer(starts_with("v_count"), names_to = "v", names_prefix = "v_count_", values_to = "v_count", values_drop_na = TRUE) %>%
  group_by(esu_dps, domain, subdomain) %>%
  mutate(v_pct = v_count / sum(v_count), v_tot = sum(v_count)) %>%
  ungroup()

df_crdcount_summ_long <-
  df_crdcount_summ %>%
  select(
    esu_dps, status, domain, subdomain, species, species2, species3,
    v_summ_pct_ag_nopasture, v_summ_pct_pasture, v_summ_pct_forest, v_summ_pct_developed, v_summ_pct_other, v_summ_pct_shrubland,
    v_summ_pct_nuts, v_summ_pct_smgrains, v_summ_pct_fruit, v_summ_pct_hay, v_summ_pct_potatoes, v_summ_pct_rice, 
    v_summ_pct_cotton, v_summ_pct_fallow, v_summ_pct_grapes, v_summ_pct_veg, v_summ_pct_corn, v_summ_pct_othercrops,
  ) %>%
  pivot_longer(starts_with("v_summ_pct"), names_to = "v", names_prefix = "v_summ_pct_", values_to = "v_summ_pct", values_drop_na = TRUE) %>%
  mutate(
    v = factor(v, levels = c("pasture", "ag_nopasture", "shrubland", "forest", "developed", "other", "nuts", "smgrains", "fruit", "hay", "potatoes", "rice", "cotton", "fallow", "grapes", "veg", "corn", "othercrops")),
    esu_dps = ordered(esu_dps, df_crdcount_summ %>% arrange(-v_summ_pct_ag) %>% pull(esu_dps)),
    domain = factor(domain, levels = c( "Puget Sound", "Interior Columbia", "Willamette/Lower Columbia", "Oregon Coast", "Southern Oregon/Northern California Coast", "North-Central California Coast", "Central Valley", "South-Central/Southern California Coast")),
    species2 = factor(species2, levels = c( "Steelhead", "Chinook", "Coho", "Sockeye", "Chum", "Pink"))
  )

```
This document summarizes land cover over the habitat range of listed salmonid populations on the West Coast, with a focus on agriculture. Included are land cover summaries for 30 Distinct Population Segments (DPS). Details on each DPS life history are also reprinted from Crozier et al. (2019), for reference, along with their map of the eight distinct recover domains (ecoregions with distinct climatic and ecological characteristics). (I've been having some trouble efficiently creating maps that include the land cover data as the raster data over the full Pacific Coast is way too large to hold in memory on my laptop.)

![](https://journals.plos.org/plosone/article/figure/image?size=large&id=10.1371/journal.pone.0217711.g001)
**Salmon recovery domains** Reprinted from Crozier et al. (2019).

Among the 30 DPS, five are listed as endangered, 23 are listed as threatened, and two are listed as species of concern. Six species are represented across 11 Recovery Domains. Total habitat area for listed DPS by species ranges from a low of 931 sq. km for Sockeye salmon (*Oncorhynchus nerka*) to a high of 375,089 sq. km for Chinook salmon (*Oncorhynchus tshawytscha*).

```{r species-table, echo=FALSE, results="asis"}
df_rd_species2 <-
  sf_recoverydomain %>%
  st_drop_geometry() %>%
  filter(species != "Not Warranted") %>%
  mutate(
    area = set_units(area, km^2),
    area_endangered = area * as.numeric(I(status == "Endangered")),
    area_threatened = area * as.numeric(I(status == "Threatened")),
    area_concern = area * as.numeric(I(status == "Species of Concern"))
  ) %>%
  group_by(species2) %>%
  summarize(
    n_domain = n_distinct(subdomain),
    n_endangered = sum(status == "Endangered"),
    n_threatened = sum(status == "Threatened"),
    n_concern = sum(status == "Species of Concern"),
    area_tot = as.vector(sum(area)),
    area_endangered = as.vector(sum(area_endangered)),
    area_threatened = as.vector(sum(area_threatened)),
    area_concern = as.vector(sum(area_concern))
  )
df_rd_species2 %>%
  rename(
    "Common Name" = species2,
    "Total DPS" = n_domain,
    "Endangered DPS" = n_endangered,
    "Threatened DPS" = n_threatened,
    "Species of Concern DPS" = n_concern,
    "Total Habitat Area (km^2)" = area_tot,
    "Endangered Habitat Area (km^2)" = area_endangered,
    "Threatened Habitat Area (km^2)" = area_threatened,
    "Species of Concern Habitat Area (km^2)" = area_concern
  ) %>%
  # mutate("All other" = )
  kable(
    digits = 0, 
    format.args = list(
      big.mark = ",", 
      scientific = FALSE
    ),
    caption = "Pacific salmonid species habitat area for listed DPS"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T, full_width = F) %>%
  # scroll_box(height = "200px") %>%
  footnote("Habitat area totals only include area for listed DPS") %>%
  print
```

# Land Cover
Land cover data is from the 2017 NASS [Cropland Data Layer](https://data.nal.usda.gov/dataset/cropscape-cropland-data-layer) (CDL). The DPS habitat range data is courtesy of Damon Holzer. Land cover shares were computed for each habitat range by extracting the raster values from the mask of the habitat range borders over the CDL.

## All DPS
The figure below presents rough summaries of land cover in the habitat ranges for all DPS. Agricultural land cover, including pasture, makes up a large share of the habitat range for many DPS. DPS in the Central Valley recovery domain have particularly large cropland land cover shares. Some Interior Columbia DPS have sizable cropland shares, driven mainly by winter wheat acreage in Eastern Washington. Upper Willamette Steelhead also have a large amount of cropland in their range, driven mostly by sod seed acreage in the Willamette Valley.

Pasture and developed land cover, both of which may also negatively affect salmon habitat quality, also make up large land cover shares of many DPS. Developed land cover is particularly common for Southern and Central California Steelhead, while pasture is especially common in the range of South-Central California Coast Steelhead.

```{r all-species, fig.height=12, fig.width=14, echo = FALSE}
# Broad Categories All Species
df_crdcount_summ_long %>%
  filter(status != "Not Warranted") %>%
  filter(!(v %in% c("nuts", "smgrains", "fruit", "hay", "potatoes", "rice", "cotton", "fallow", "veg", "grapes", "corn", "othercrops"))) %>%  #Drop the specific categories to see the broad categories
#  filter(v != "forest", v != "developed", v != "other") %>%  #toggle to see just cropland and pasture on the figure
ggplot() +
  geom_col(aes(x = esu_dps, y = v_summ_pct, fill = v), position = "stack", width = 0.75) +
  geom_point(aes(x = esu_dps, y = -0.05, color = status), fill = NA, size = 7, shape = "square", position = "identity") +
  geom_text(aes(x = esu_dps, y = 1.25, label = domain), vjust = 0.5, size = 3.5, fontface = "plain") +
  scale_fill_manual(
    limits = c("ag_nopasture", "pasture", "shrubland", "forest", "developed", "other"),
    labels = c("Cropland", "Pasture", "Shrubland", "Forest", "Developed", "Other"),
    values = c("darkgoldenrod", "goldenrod", "springgreen3", "darkgreen", "darkgrey", "lightgrey"),
    name = "Land Cover"
  ) +
  scale_color_manual(
    limits = c("Not Warranted", "Species of Concern", "Threatened", "Endangered"),
    values = c("green", "yellow", "darkorange", "red"),
    name = "Status"
  ) +
  scale_y_continuous(breaks = c(-0.05, 0.5, 1.25), labels = c("Status", "Land Use", "Domain"), limits = c(-0.05, 1.4),
                     minor_breaks = c(0, 0.25, 0.5, 0.75, 1),
                     sec.axis = dup_axis(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = scales::label_percent())) +
  guides(
    x = guide_axis(position = "top", order = 0, check.overlap = TRUE),
    x.sec = guide_axis(position = "bottom", order = 1, check.overlap = TRUE),
    y = guide_axis(position = "left")
  ) +
  theme(
    axis.text.y = element_text(hjust = 1, vjust = 0.5),
    panel.spacing = unit(0.1, "lines"),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.title = element_blank(),
    panel.grid.minor.x = element_line(linetype = "dashed", color = "black"),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.background = element_blank(),
    # panel.background = element_rect(color = "black"),
    strip.background = element_rect(fill = "lightgrey"),
    strip.placement = "outside"
  ) +
  facet_grid(rows = "species2", switch = "y", scales = "free_y", space = "free_y") +
  coord_flip() +
  ggtitle("Land cover in habitat range for listed salmonid DPS")

# Only agricultural use categories, all species All Species
df_crdcount_summ_long %>%
  filter(status != "Not Warranted") %>%
  filter(!(v %in% c("pasture", "ag_nopasture","shrubland", "forest", "developed", "other"))) %>%
ggplot() +
  geom_col(aes(x = esu_dps, y = v_summ_pct, fill = v), position = "stack", width = 0.75) +
  geom_point(aes(x = esu_dps, y = -0.05, color = status), fill = NA, size = 7, shape = "square", position = "identity") +
  geom_text(aes(x = esu_dps, y = 1.25, label = domain), vjust = 0.5, size = 3.5, fontface = "plain") +
  scale_fill_manual(
    limits = c( "nuts", "smgrains", "fruit", "hay", "potatoes", "rice", "cotton", "fallow", "grapes", "veg", "corn", "othercrops"),
    labels = c( "Nuts", "Small grains", "Fruit", "Hay", "Potatoes", "Rice", "Cotton", "Fallow cropland", "Grapes", "Vegetables", "Corn", "Other crops"),
    values = c( "darkgoldenrod", "goldenrod", "darkred", "darkgreen", "black",  "violet","pink", "lightgrey", "purple", "lightgreen", "gold", "lightblue"),
    name = "Agricultural Land Cover"
  ) +
  scale_color_manual(
    limits = c("Not Warranted", "Species of Concern", "Threatened", "Endangered"),
    values = c("green", "yellow", "darkorange", "red"),
    name = "Status"
  ) +
  scale_y_continuous(breaks = c(-0.05, 0.5, 1.25), labels = c("Status", "Land Use", "Domain"), limits = c(-0.05, 1.4),
                     minor_breaks = c(0, 0.25, 0.5, 0.75, 1),
                     sec.axis = dup_axis(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = scales::label_percent())) +
  guides(
    x = guide_axis(position = "top", order = 0, check.overlap = TRUE),
    x.sec = guide_axis(position = "bottom", order = 1, check.overlap = TRUE),
    y = guide_axis(position = "left")
  ) +
  theme(
    axis.text.y = element_text(hjust = 1, vjust = 0.5),
    panel.spacing = unit(0.1, "lines"),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.title = element_blank(),
    panel.grid.minor.x = element_line(linetype = "dashed", color = "black"),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.background = element_blank(),
    # panel.background = element_rect(color = "black"),
    strip.background = element_rect(fill = "lightgrey"),
    strip.placement = "outside"
  ) +
  facet_grid(rows = "species2", switch = "y", scales = "free_y", space = "free_y") +
  coord_flip() +
  ggtitle("Agricultural land use in habitat range for listed salmonid DPS")
```

## DPS Breakdowns
This section provides breakdowns of the most common land covers for all 30 DPS, along with a rough map of their range. The tables provided reveal the most common crops grown in the ranges of each DPS (if any). For species in the Central Valley, rice, walnuts, and almonds are the most common crops in their habitat range. All three crops have high water demands. Winter wheat is the most common crop in the Interior Columbia ranges. Note that winter wheat is commonly rotated against fallow (i.e. one year wheat, the next year fallow, for each field), which is also very common in these ranges. If conservation tillage is not used, fallow fields can be particularly vulnerable to sediment and nutrient run-off.

```{r species-summs, fig.height=12, fig.width=6, results="asis", echo=FALSE}
# Load states borders and transform to match crs
states_core <- st_as_sf(maps::map("state", regions = c("california", "oregon", "washington", "idaho"), plot = FALSE, fill = TRUE)) %>% st_transform(st_crs(sf_recoverydomain))
states_expand <- st_as_sf(maps::map("state", regions = c("california", "oregon", "washington", "idaho", "montana", "wyoming", "arizona", "nevada", "utah", "colorado", "new mexico"), plot = FALSE, fill = TRUE)) %>% st_transform(st_crs(sf_recoverydomain))
land <- st_as_sf(maps::map("world", regions = c("Canada", "Mexico"), plot = FALSE, fill = TRUE)) %>% st_transform(st_crs(sf_recoverydomain))

# List of distinct population segments
dps <- df_crdcount %>% distinct(esu_dps) %>% arrange(esu_dps) %>% pull(esu_dps)

# Loop profile over all DPS...
for(i in dps) {
  # Print species title and some info
  cat("  \n###", i)
  cat("  \n**Status:**", df_crdcount %>% filter(esu_dps == i) %>% pull(status))
  cat("  \n**Domain:**", df_crdcount %>% filter(esu_dps == i) %>% pull(domain))
  cat("  \n")

  # Build range map
  p_range <-
    ggplot() +
    # Basic landmass polygon
    geom_sf(data = land, fill = "antiquewhite1") +
    # State borders polygon
    geom_sf(data = states_expand, fill = "antiquewhite1") +
    # Highlight the range
    geom_sf(data = sf_recoverydomain %>% filter(esu_dps == i), fill = NA, color = "red") +
    # Set bounding box
    coord_sf(
      xlim = st_bbox(states_core)[c(1,3)],
      ylim = st_bbox(states_core)[c(2,4)]
    ) +
    # Some themeing
    theme_void() +
    theme(
      panel.background = element_rect(fill = "aliceblue", color = NA)
    )
  
  # Build land cover bar plot
  p_cover <-
    df_crdcount_summ_long %>%
    # Isolate just the one species
    filter(esu_dps == i) %>%
    ggplot() +
    # Plot the land cover breakdown
    geom_col(aes(x = esu_dps, y = v_summ_pct, fill = v), position = "stack", width = 0.75) +
    # Set the color scale for land cover breakdown
    scale_fill_manual(
      limits = c("ag_nopasture", "pasture", "shrubland", "forest", "developed", "other"),
      labels = c("Cropland", "Pasture", "Shrubland", "Forest", "Developed", "Other"),
      values = c("darkgoldenrod", "goldenrod", "palegreen", "darkgreen", "darkgrey", "lightgrey"),
      name = ""
    ) +
    # Set up the y-axis
    scale_y_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = scales::label_percent()) +
    # guides(
    #   x = guide_axis(position = "top", order = 0, check.overlap = TRUE),
    #   y = guide_axis(position = "left")
    # ) +
    # Some themeing
    theme_void() +
    theme(
      # axis.text.x = element_text(hjust = 1, vjust = 0.5),
      panel.spacing = unit(0.1, "lines"),
      axis.line = element_blank(),
      # axis.ticks = element_blank(),
      axis.text.x = element_text(),
      axis.title = element_blank(),
      panel.grid.major.x = element_line(linetype = "dashed", color = "black"),
      legend.position = "bottom",
      legend.box = "horizontal",
      legend.background = element_blank(),
      plot.title.position = "plot",
      # panel.background = element_rect(color = "black"),
      strip.background = element_rect(fill = "lightgrey"),
      strip.placement = "outside"
    ) +
    # Flip so it fits under the map
    coord_flip() +
    # Add title
    ggtitle("Land Cover Share")
  # p_row <- plot_grid(p_range, p_cover, rel_heights = c(0.8, 0.2))
  
  #Plot the agricultural land uses inside each area
   (border <- sf_recoverydomain %>% 
      arrange(area) %>% 
      filter(esu_dps == i)) 
  clip1 <- raster::crop(raster_cdl_merge, extent(border)) # Clip cdl to rectangle extents of the polygon
  clip2 <- mask(clip1, border) # Mask cdl to only what's within polygon
   (df_clip2 <- as.data.frame(clip2, xy = TRUE))
   df_clip2 <- df_clip2 %>% 
     mutate(cdl_west = updateNamesCDL(cdl_west))

# Create table of counts by land use type (proportion doesn't work because we have all those NAs)
# This is also used as the substitution matrix to group all land use types into crop categories
 types <- df_clip2  %>%
   group_by(cdl_west)  %>%
   count
 
 types <- types %>%
  mutate(cropgroup = if_else(str_detect(cdl_west, "Barley|Rye|Wheat|Oats|Triticale|Sorghum|Buckwheat|Millet|Dbl"), "smgrains",
                    if_else(str_detect(cdl_west, "Walnuts|Almonds|Pecans|Pistachios"), "nuts" ,
                    if_else(str_detect(cdl_west, "Apples  | Apricots|berries|  Cantaloupes|Cherries|Citrus|Watermelons|Melons|Nectarines|Oranges|Peaches|Pears|Plums|Pomegranates|Other Tree|Olives"), "fruits",
                    if_else(str_detect(cdl_west, "Tomatoes|Asparagus|Broccoli|Cabbage|Carrots|Cauliflower|Cucumbers|Garlic|Gourds|Greens|Lettuce|Onions|Peas|Peppers|Popcorn|Pumpkins|Radish|Squash|Sweet Corn|Sweet Potatoes|Turnips"), "veg",
                    if_else(str_detect(cdl_west,  "Chick Peas|Christmas Trees|Dry Beans|Flaxseed|Herbs|Hops|Lentils|Mint|Misc Vegs|Mustard|Other Crops|Rape|Safflower|Sod/Grass Seed|Sugarbeets|Sunflower|Vetch"), "other_crops",
                    if_else(str_detect(cdl_west,  "Alfalfa|Other Hay"), "hay",
                    if_else(str_detect(cdl_west,  "Rice"), "rice",
                    if_else(str_detect(cdl_west,  "Cotton"), "cotton",
                    if_else(str_detect(cdl_west,  "Fallow"), "fallow",
                    if_else(str_detect(cdl_west,  "Grapes"), "grapes",
                    if_else(str_detect(cdl_west,  "Pasture"), "pasture",
                    "other")))))))))))) 
types$cropgroup <- if_else(types$cdl_west == "Potatoes", "potatoes", types$cropgroup)
types$cropgroup <- if_else(types$cdl_west == "Corn", "corn", types$cropgroup)
types$cropgroup <- if_else(types$cdl_west == "Open Water", "Open Water", types$cropgroup)


# Join types to the land use data frame
df_clip2_types = left_join(df_clip2, types)
# Plot crop categories (types)
p_ag_cover <-
ggplot() +
  geom_raster(data = df_clip2_types, aes(x = x, y = y, fill = cropgroup)) +
  scale_fill_manual(
    limits = c( "nuts", "smgrains", "fruit", "hay", "potatoes", "rice", "cotton", "fallow", "grapes", "veg", "corn", "othercrops", "Open Water", "Pasture", "other"),
    labels = c( "Nuts", "Small grains", "Fruit", "Hay", "Potatoes", "Rice", "Cotton", "Fallow cropland", "Grapes", "Vegetables", "Corn", "Other crops", "Open Water", "Pasture", "other"),
    values = c( "darkgoldenrod", "goldenrod", "darkred", "darkgreen", "black",  "violet","pink", "grey", "purple", "lightgreen", "gold", "lightblue", "darkblue", "springgreen3", "antiquewhite1"))

  
  # Build titles as seperate objects so we can put them in the composite figure
  p_title <-
    ggdraw() + 
    draw_label(
      "Habitat range and land cover share",
      fontface = 'bold',
      x = 0,
      hjust = 0
    ) +
    theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0, 0)
    )
  p_subtitle <-
    ggdraw() + 
    draw_label(
      i,
      fontface = 'bold',
      size = 12,
      x = 0,
      hjust = 0
    ) +
    theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0, 0)
    )
  
  # Plot it all
  # Some subtitles
  cat("  \n#### Habitat range and land cover")  
  cat("  \n")
  
  # Plot the composite figure
  plot_grid(
    p_range, p_cover, p_ag_cover,
    ncol = 1,
    # rel_heights values control vertical title margins
    rel_heights = c(0.7, 0.15)
  ) %>%
    print

  # Print the table of land cover shares (with finer crop breakdown)
  df_crdcount_long %>%
    # Grab just the dps we're interested in
    filter(esu_dps == i) %>%
    # Select just the value columns
    select(v, v_pct) %>% 
    # Arrange so the largest share is on top
    arrange(-v_pct) %>%
    # Drop shares with less than 1% of land cover
    filter(v_pct >= 0.01) %>%
    # Clean up so it's pretty
    mutate(
      v = make_clean_names(v, "sentence"),
      v_pct = scales::percent(v_pct, accuracy = 0.1)
    ) %>%
    # Give the variables descriptive pretty labels for column headers
    rename(
      "Land cover" = v,
      "%" = v_pct
    ) %>%
    # mutate("All other" = )
    # Transform into html table
    kable() %>%
    # Style it
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), fixed_thead = T, full_width = F, position = "float_right") %>%
    # scroll_box(height = "200px") %>%
    # Add footnote on where the other land covers went
    footnote("All other land covers < 1%") %>%
    # Make sure it shows up
    print
    # Add one more linebreak so we don't bleed into the next one report
    cat("  \n")
}
```
# Life Histories and Seasonality
The effects of agriculture on salmon will depend on their life histories. DPS migrate, rear, and spawn at different times throughout the year. When environmentally stressful agricultural practices occurs also varies by crop.

To facilitate discussion about possible temporal overlaps, I provide the life histories of each species below, followed by links to the crop progress reports for commodity crops in the four states. Reports for crops with major geographic overlap with specific species can be compared to their life histories to reveal when certain species might be particularly vulnurable to field operations.

Note that the [supplementary material](https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0217711.s007&type=supplementary) in Crozier et al. (2019) has maps that break DPS ranges down by use type (i.e. migration, rearing, spawning), which could allow us to break down land cover in even more detail over each species' life history.

![](https://journals.plos.org/plosone/article/figure/image?size=large&id=10.1371/journal.pone.0217711.g002)
**Salmon life histories** Reprinted from Crozier et al. (2019). Colors represent life stages, where yellow indicates adult freshwater migration and holding, red spawning, orange incubation, light blue juvenile freshwater rearing and migration, green estuary and nearshore rearing, and dark blue marine stage.

[NASS Crop Progress Reports](https://www.nass.usda.gov/Charts_and_Maps/Crop_Progress_&_Condition/2019/index.php)  
[CA](https://www.nass.usda.gov/Charts_and_Maps/Crop_Progress_&_Condition/2019/CA_2019.pdf) | [OR](https://www.nass.usda.gov/Charts_and_Maps/Crop_Progress_&_Condition/2019/OR_2019.pdf) | [WA](https://www.nass.usda.gov/Charts_and_Maps/Crop_Progress_&_Condition/2019/WA_2019.pdf) | [ID](https://www.nass.usda.gov/Charts_and_Maps/Crop_Progress_&_Condition/2019/ID_2019.pdf)

# Citations
Crozier, L. G., McClure, M. M., Beechie, T., Bograd, S. J., Boughton, D. A., Carr, M., ... & Hazen, E. L. (2019). Climate vulnerability assessment for Pacific salmon and steelhead in the California Current Large Marine Ecosystem. *PloS One*, 14(7), e0217711. [Link](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0217711)

